---
title: "Understanding Data Distributions Recitation Solutions ☕"
author: "Jessica Cooperstone"
subtitle: "Week 7"
format:
  html:
    toc: true
    toc-depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Today you will be investigating some [data](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md) from the Coffee Quality Database on coffee evaluation by the Coffee Quality Institute. You will look at the distribution of coffee evaluation scores across different characteristics, and include various metadata.

More information can be found on the Tidy Tuesday github repo on [coffee ratings](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md).

### Read in data
```{r load data}
tuesdata <- tidytuesdayR::tt_load('2020-07-07')
coffee_ratings <- tuesdata$coffee_ratings
```

### Load libraries
```{r libraries, warning = FALSE, message = FALSE}
library(tidyverse) # for everything
library(ggridges) # for ridgeline plots
library(ggdist) # for nice dotplots
```

## Total cupping score in arabica and robusta
Make 3 different visualizations that shows the distribution of total cupping score across arabica and robusta beans. Make the plots so you think they look good.

### A histogram
Since there are so few robusta observations, I decided to make the y-axes on different scales
```{r}
coffee_ratings %>%
  ggplot(aes(x = total_cup_points, fill = species)) +
  geom_histogram(bins = 150) +
  geom_vline(aes(xintercept = mean(total_cup_points)), color = "black") +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  coord_cartesian(xlim = c(60,100)) +
  facet_wrap(vars(species),
             nrow = 2, # make two rows so can align histograms top to bottom
             scales = "free_y",
             strip.position = "top") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Total cupping score (out of 100)",
       y = "",
       fill = "Species",
       title = "Distribution of cupping scores across 14,000 coffee samples",
       subtitle = "Note the y-axes are different between plots",
       caption = "Vertical line represents the median total cupping score across all samples")
```

### Density plot
I might like this a bit better than a histogram.
```{r}
coffee_ratings %>%
  ggplot(aes(x = total_cup_points, fill = species)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(total_cup_points)), color = "black") +
  coord_cartesian(xlim = c(65,92)) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  facet_wrap(vars(species),
             nrow = 2, # make two rows so can align histograms top to bottom
             scales = "free_y") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Total cupping score (out of 100)",
       y = "",
       fill = "Species",
       title = "Distribution of cupping scores across 14,000 coffee samples",
       caption = "Vertical line represents the median total cupping score across all samples")
```

### Dot plot
You can also see here how many fewer robusta observations there are.
```{r}
# dot plot
coffee_ratings %>%
  ggplot(aes(x = total_cup_points, color = species, fill = species)) +
  geom_dots() +
  scale_color_manual(values = c("#C19A6B", "#5C4033")) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  theme_ggdist() +
  xlim(50,100) +
  theme(legend.position = c(.18, .99),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.box.background = element_rect(size = 0.5),
        legend.box.margin = margin(5, 5, 5, 5)) +
  labs(x = "Total cupping score (out of 100)",
       y = "",
       color = "Species",
       fill = "Species")
```

### Ridgeline plot

```{r}
# ridgeline plot
coffee_ratings %>%
  ggplot(aes(x = total_cup_points, y = species, fill = species)) +
  stat_density_ridges(quantile_lines = TRUE,
                      quantiles = 2,
                      alpha = 0.5) +
  xlim(55, 100) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  theme_ggdist() +
  theme(legend.position = "none") +
  labs(x = "Total cupping score (out of 100)",
       y = "",
       title = "Distribution of cupping scores across 14,000 coffee samples",
       subtitle = "The brown color theme is very coffee-esque")
```

## Individual characteristic cupping scores in arabica and robusta
Make 3 different visualizations that show the distribution of all the individual contributors to total cupping score across arabica and robusta in one plot.

Wrangling:
```{r}
# go from wide (each coffee attribute in a separate column)
# to long data (1 column w/ all characteristics, 1 column w/ all ratings)
coffee_ratings_tidy <- coffee_ratings %>%
  pivot_longer(cols = aroma:cupper_points, # columns from aroma to cupper_points
               names_to = "characteristic",
               values_to = "rating")
```

Prepare to clean up facet strip text using the function `labeller()`.
```{r}
# getting labels ready for plotting
# what are the coffee characteristics again?
(coffee_characteristics <- unique(coffee_ratings_tidy$characteristic) %>%
  sort()) # sort alphabetically, arrange won't work here bc not numeric

# create a vector of the coffee characteristic names as i want them to appear on the plot
coffee_labels <- c("Acidity",
                   "Aftertaste",
                   "Aroma",
                   "Balance",
                   "Body",
                   "Clean Cup",
                   "Cupper Points",
                   "Flavor",
                   "Sweetness",
                   "Uniformity")

# tell coffee_labels which original label to refer to
# these need to be in the same order (which is why i used sort())
names(coffee_labels) <- coffee_characteristics
```

Managing fonts:
```{r}
# get fonts not default available in R
library(sysfonts) # aux packagew here fonts live
library(showtext) # package that helps use non-standard fonts
library(ragg)

# add the font Atkison Hyperlegible bc i like it
font_add_google("Atkinson Hyperlegible")

# what fonts do i have to choose from?
# remove head() to see them all
head(font_info_google())

# use to indicate that showtext is needed 
showtext_auto()
```


### Boxplots
This is just ok.
```{r}
coffee_ratings_tidy %>%
  ggplot(aes(x = species, y = rating, fill = species)) +
  geom_boxplot(color = "black", alpha = 0.8) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  coord_cartesian(ylim = c(5,10)) +
  facet_wrap(vars(characteristic),
             labeller = labeller(characteristic = coffee_labels), # all that work we did earlier
             nrow = 2) + # 2 rows in the faceted plot
  theme_classic() + 
  theme(legend.position = "none",
        text = element_text(family = "Atkinson Hyperlegible")) + # changing font
  labs(x = "Coffee Species",
       y = "Cupper rating (out of 10)",
       title = "Distribution of cupper scores for individual coffee attributes",
       caption = "Line represents the median rating per species")
```

### Violin plot
```{r, warning = FALSE, message = FALSE}
# violin plot
coffee_ratings_tidy %>%
  ggplot(aes(x = species, y = rating, fill = species)) +
  geom_violin(draw_quantiles = 0.5, color = "black", alpha = 0.8) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  coord_cartesian(ylim = c(5,10)) +
  facet_wrap(vars(characteristic),
             labeller = labeller(characteristic = coffee_labels), # all that work we did earlier
             nrow = 2) + # 2 rows in the faceted plot
  theme_classic() + 
  theme(legend.position = "none",
        text = element_text(family = "Atkinson Hyperlegible")) + # changing font
  labs(x = "Coffee Species",
       y = "Cupper rating (out of 10)",
       title = "Distribution of cupper scores for individual coffee attributes",
       caption = "Line represents the median rating per species")
```

### Dot plots
This is just ok.
```{r}
coffee_ratings_tidy %>%
  ggplot(aes(x = species, y = rating, color = species)) +
  geom_dots(side = "both", layout = "swarm") +
  scale_color_manual(values = c("#C19A6B", "#5C4033")) +
  coord_cartesian(ylim = c(6,10)) +
  facet_wrap(vars(characteristic),
             labeller = labeller(characteristic = coffee_labels),
             nrow = 2) +
  theme_classic() + 
  theme(legend.position = "none",
        text = element_text(family = "Atkinson Hyperlegible")) + # changing font
  labs(x = "Cupper rating (out of 10)",
       y = "Coffee Species",
       title = "Distribution of cupper scores for individual coffee attributes")
```

### Density plot
This one I think is my favorite.
```{r, warning = FALSE, message = FALSE}
coffee_ratings_tidy %>%
  ggplot(aes(x = rating, y = species, fill = species)) +
  geom_density_ridges(alpha = 0.8,
                      quantile_lines = TRUE,
                      quantiles = 2) +
  scale_fill_manual(values = c("#C19A6B", "#5C4033")) +
  coord_cartesian(xlim = c(6,10)) +
  facet_wrap(vars(characteristic),
             labeller = labeller(characteristic = coffee_labels),
             nrow = 2) +
  theme_classic() + 
  theme(legend.position = "none") +
  labs(x = "Cupper rating (out of 10)",
       y = "Coffee Species",
       title = "Distribution of cupper scores for individual coffee attributes")

```


