---
title: "Syllabus, Autumn 2025"
format:
  html:
    toc: true
    toc-depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Instructor

**Instructor:** [Jessica Cooperstone, Ph.D.](https://www.cooperstonelab.com/)\
**Email address:** cooperstone dot 1 at osu dot edu (preferred contact method)\
**Phone number:** 614-292-2843 (non-preferred contact method)\
**TA**: [Daniel Quiroz Moreno](https://danielquiroz97.gitlab.io/), quirozmoreno dot 1 at osu dot edu\
**Office hours:** TBD

<!-- - Jess: Tuesdays 1-2pm Howlett 348 or via [Zoom](https://go.osu.edu/dataviz-zoom) -->

<!-- - Daniel: Thursdays 2-3pm Howlett 340 or via [Zoom](https://go.osu.edu/dataviz-zoom) (tell Daniel by the start of office hours if you're planning to attend) -->

## Course meeting time and place

Tuesdays **synchronously** from 4:10-5:55 pm in:

-   Howlett Hall 116 (Columbus)
-   Williams Hall 123 (Wooster)
-   Virtually via Zoom <https://go.osu.edu/dataviz-zoom>

## Course description

This course aims to introduce students to the principles and practice of data visualization. Students will learn fundamental principles of data visualization and create figures that appropriately and ethically represent their data. Data visualizations will be created in the R programming environment, using tools including the grammar of graphics implemented in ggplot2. In the process of creating visualizations, students will also become familiar with data handling and wrangling in R.

**Course learning outcomes**\
By the end of this course, students should successfully be able to:

-   Recall and describe the fundamental goals and principles of data visualization.
-   Distinguish between good and bad visualizations, and understand how to make those are ineffective more effective.
-   Learn to use R, R Markdown, and ggplot to make clear, descriptive, and aesthetic visualization.
-   Apply principles learned in class, both theoretical and technical, to create effective visualizations.

## Course delivery

**Mode of delivery:** This course is a hybrid-delivered course where I provide \~ 50 min of lecture material, followed by \~1 hour of recitation activities to do in-class (with assistance from the instructor and the TA). You can attend class in person or virtually via Zoom. You are also welcome to come to class with questions about the weekly videos, or problems you are currently encountering in creating your visualizations.

**Attendance and participation requirements:** This is a hybrid course, but is taught synchronously, meaning it is expected that you attend class, either in person or virtually (via Zoom) during its meeting time. I will not take attendance. If circumstances require you to miss class, it will be expected you watch the recorded sessions, and catch up on material on your own. I have found that students who attend classes more quickly and completely master course content.

**Class recordings:** To help you master material, and to better accommodate students, classes will be recorded, and recordings uploaded directly after class to Carmen. You can find a link to a OneDrive folder with the recorded lectures on the Syllabus page on Carmen.

## Course schedule

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(readxl)
library(knitr)
library(tidyverse)
schedule <- read_excel("../files/schedule_modules.xlsx",
                       sheet = "Schedule_date_2025") 
```

```{r, echo = FALSE, fig.width = 6}
kable(schedule)
```

::: {.callout-note title="Non-standard days"}
There is no class on November 11 for Veterans Day. On November 25 (the short week of Thanksgiving), I will provide to you the lecture material asynchronously and you can go through the recitation material on your own. Daniel and I will be available to answer questions during office hours, or at the beginning of the next class period.
:::

## Course resources

There are no required textbooks for this course, though you will find many of the recommend texts and resources belowvery useful.

-   [R for Data Science](https://r4ds.had.co.nz/index.html) by Hadley Wickham and Garrett Grolemund
-   [Introduction to Data Science](https://rafalab.github.io/dsbook/) by Rafael Irizarry
-   [Data Visualization with R](https://rkabacoff.github.io/datavis/) by Rob Kabacoff
-   [Data Visualization](https://socviz.co/), A practical introduction by Kieran Healy
-   [Hands-On Data Visualization](https://handsondataviz.org/) by Jack Dougherty and Ilya Ilyankou
-   [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/) by Claus O. Wilke
-   [ggplot2, Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html) by Hadley Wickham, Danielle Navarro, and Thomas Lin Pedersen
-   [Modern Data Science with R](https://mdsr-book.github.io/mdsr2e/) by Benjamin S. Baumer, Daniel T. Kaplan, and Nicholas J. Horton
-   [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/) by Yihui Xie, Christophe Dervieux, Emily Riederer
-   [The tidyverse style guide](https://style.tidyverse.org/)
-   [A ggplot2 tutorial for beautiful plotting in R](https://cedricscherer.netlify.app/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/#toc) by Cédric Scherer
-   [RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/), including [`ggplot2`](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf), [dplyr](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf), [tidyr](https://raw.githubusercontent.com/rstudio/cheatsheets/main/tidyr.pdf), [readr and other data import packages](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-import.pdf), [stringr for managing character strings](https://raw.githubusercontent.com/rstudio/cheatsheets/main/strings.pdf), [forcats for managing factors](https://raw.githubusercontent.com/rstudio/cheatsheets/main/factors.pdf), and [R Markdown](https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf)

## Required software

-   R: We will use the R programming environment for this class <https://www.r-project.org/> (free). You can do so many things in R (including building this course website).
-   RStudio Desktop: This IDE (integrated development environment) allows a user-friendly interface with the R programming environment, which we will use in class as well. You must have R before you download [RStudio](https://www.rstudio.com/products/rstudio/download/) (free).
-   Microsoft Office 365: All Ohio State students are now eligible for free Microsoft Office 365. Full instructions for downloading and installation can be found at [go.osu.edu/office365help](www.go.osu.edu/office365help).

## Prior R experience

You do not need to be an R expert for this class, but I will assume working-level knowledge of R programming. If you have no experience with R, but would still like to take this class, you can. I ask then you get yourself up to speed by taking this free online class <https://www.edx.org/course/data-science-r-basics> (audit only) before the start of the 3rd week of class. The course will take 8-16 hours to complete so please leave yourself enough time to do so before week 3. Tips and tricks in R will be scattered throughout the course material.

## Grading

### How your grade is calculated

| Assignment category (how many assignments) | Points |
|----|----|
| Module assignments (4) | 32 (8 points per assignment) |
| Recitations (8, you pick which ones) | 8 (1 point per recitation submitted) |
| Class reflections (10, you pick which ones) | 20 (2 points per reflection) |
| Capstone assignment (1 plan, 1 assignment) | 10 for the capstone plan, 30 for the assignment |


See the Assignments tab for additional information.

### Assignment descriptions

#### Module assignments

*Description*: After each module, there will be an assignment to provide practice for the techniques learned in class. Assignments will be posted at least one week prior to their due date, and due dates can be found on Carmen.

*Grading*: Each part of the assignment will have a certain number of points associated with it, provided along with the assignment.

*Academic Integrity*: Students may use class notes and class resource materials. Students are not permitted to collaborate on assignments. Students must complete work on their own.

####  Recitation submissions

*Description*: I ask you submit 8 of 11 recitations to Carmen to show you have made a good faith effort to engage with the course material. 

*Grading*: I will mark these at 0 or 1 points, with 1 point given for completion of at least 70% of the assignment.

*Academic Integrity*: Students may use class notes and class resource materials. Students are not permitted to collaborate on assignments. Students must complete work on their own. Generative AI is not to be used. 


#### Class reflections

After each week, you will write a 1 paragraph reflection on the material that was presented in class. This can include your thoughts on how you will use these lessons in your own research and data visualizations, ways in which you have investigated this topic (or expect to) on your own, or what else you’d like to learn in this area. The purpose of this assignment is not to be burdensome, but to keep you engaged in the course material, and providing feedback to me on what parts you’ve found useful, what you’ve struggled with, and what you’d like to see more of in the future.

There will be 10 class reflections (you can select which classes you want to reflect upon).

**Due Date:** Reflections are due 1 week after each class. For example, if class is on Tuesday September 1, the reflection for that class is due on Tuesday September 8 by 11:59pm.

*Grading*: Reflections will be graded as follows:

-   Full credit: Reflections thoughtfully engage with course content, demonstrate the student thought about material and how it would (or wouldn’t) be relevant to their work and development. This is the level of engagement I expect for this course.
-   Half credit: Reflections are superficial and demonstrate minimal engagement with the course content. A half credit grade indicates better engagement is required for the next reflection. I do not expect to assign these grades often.

*Academic Integrity*: Students may use class notes and class resource materials. Students are not permitted to collaborate on assignments. Students must complete work on their own.

#### Capstone assignment

*Description*: At the end of the semester, you will complete a capstone assignment where you create a series of visualizations based on your research data, data coming from your lab, or other data that is publicly available. I expect this assignment to be completed in R Markdown, annotated, and knitted into an easy-to-read .html file. I also expect your code to be fully commented such that I can understand what you are doing with each step, and why. You will be required to submit a capstone assignment “plan” by the beginning of November.

*Grading*: Guidance will be provided for the grading of the capstone assignment. Assignments completed outside R Markdown, or not knitted to a .html are not acceptable.

*Academic Integrity*: Students may use class notes and class resource materials. Students are not permitted to collaborate on assignments. Students must complete work on their own.

#### Due dates

```{r, echo = FALSE}
library(readxl)
library(knitr)
due_dates <- read_excel("../files/schedule_modules.xlsx",
                       sheet = "module_due_dates_2025")
```

```{r, echo = FALSE, fig.width = 6}
kable(due_dates)
```

### Late assignments

I expect you will turn assignments in on time. Late assignments are not accepted. If there are extenuating circumstances that prevent you from turning in an assignment on time, please connect with me as soon as possible after such a situation arises for discussion about a possible deadline extension.

### Grading scale

```{r, echo = FALSE}
library(readxl)
library(knitr)
grading_scale <- read_excel("../files/grading_scale.xlsx",
                       sheet = "Sheet1")
```

```{r, echo = FALSE, fig.width = 4}
kable(grading_scale)
```

### Instructor feedback and response time

Your instructional team is here to help you proceed through this course. We will match but not exceed the enthusiasm that you give to learn the course content. 

I am providing the following list to give you an idea of my intended availability throughout the course. (Remember that you can call 614-688-4357(HELP) at any time if you have a technical problem.)

-  Grading and feedback: For assignments, you can generally expect feedback within 7 days.
-  Email: We will reply to emails within 48 hours on days when class is in session at the university.


## Other course policies

### Discussion and communication guidelines

I expect all communication will be respectful and thoughtful.

### Academic Misconduct/Academic Integrity

See Descriptions of major course assignments, above, for my specific guidelines about collaboration and academic integrity in the context of this online class.
Academic integrity is essential to maintaining an environment that fosters excellence in teaching, research, and other educational and scholarly activities. Thus, The Ohio State University and the Committee on Academic Misconduct (COAM) expect that all students have read and understand the University's Code of Student Conduct, and that all students will complete all academic and scholarly assignments with fairness and honesty. Students must recognize that failure to follow the rules and guidelines established in the University's Code of Student Conduct and this syllabus may constitute Academic Misconduct.

The Ohio State University's Code of Student Conduct (Section 3335-23-04) defines academic misconduct as: Any activity that tends to compromise the academic integrity of the University or subvert the educational process. Examples of academic misconduct include (but are not limited to) plagiarism, collusion (unauthorized collaboration), copying the work of another student, and possession of unauthorized materials during an examination. Ignorance of the University's Code of Student Conduct is never considered an excuse for academic misconduct, so I recommend that you review the Code of Student Conduct and, specifically, the sections dealing with academic misconduct.

If I suspect that a student has committed academic misconduct in this course, I am obligated by University Rules to report my suspicions to the Committee on Academic Misconduct. If COAM determines that you have violated the University's Code of Student Conduct (i.e., committed academic misconduct), the sanctions for the misconduct could include a failing grade in this course and suspension or dismissal from the University.

**If you have any questions about the above policy or what constitutes academic misconduct in this course, please contact me.**

### Use of Generative AI (e.g., ChatGPT, Co-Pilot or similar)

Artificial intelligence tools, like ChatGPT, Microsoft Co-Pilot or similar are not to be used on assignments unless instructions explicitly indicate to do so. Suspected unauthorized use of generative AI on assignments of any type will be reported to COAM.

### Creating an environment free from harassment, discrimination, and sexual misconduct

The Ohio State University is committed to building and maintaining a community to reflect diversity and to improve opportunities for all. All Buckeyes have the right to be free from harassment, discrimination, and sexual misconduct. Ohio State does not discriminate on the basis of age, ancestry, color, disability, ethnicity, gender, gender identity or expression, genetic information, HIV/AIDS status, military status, national origin, pregnancy (childbirth, false pregnancy, termination of pregnancy, or recovery therefrom), race, religion, sex, sexual orientation, or protected veteran status, or any other bases under the law, in its activities, academic programs, admission, and employment. Members of the university community also have the right to be free from all forms of sexual misconduct: sexual harassment, sexual assault, relationship violence, stalking, and sexual exploitation.

To report harassment, discrimination, sexual misconduct, or retaliation and/or seek confidential and non-confidential resources and supportive measures, contact the Civil Rights Compliance Office:

Online reporting form at civilrights.osu.edu
Call 614-247-5838 or TTY 614-688-8605,
Or Email civilrights@osu.edu

The university is committed to stopping sexual misconduct, preventing its recurrence, eliminating any hostile environment, and remedying its discriminatory effects. All university employees have reporting responsibilities to the Civil Rights Compliance Office to ensure the university can take appropriate action:

-  All university employees, except those exempted by legal privilege of confidentiality or expressly identified as a confidential reporter, have an obligation to report incidents of sexual assault immediately.
-  The following employees have an obligation to report all other forms of sexual misconduct as soon as practicable but at most within five workdays of becoming aware of such information: 1. Any human resource professional (HRP); 2. Anyone who supervises faculty, staff, students, or volunteers; 3. Chair/director; and 4. Faculty member.

In addition, this course adheres to The Principles of Community adopted by the College of Food, Agricultural, and Environmental Sciences. These principles can be found at https://cfaes.osu.edu/about/cfaes-principles-community. If you have been a victim of or a witness to harassment or discrimination or a bias incident, you can report it online and anonymously (if you choose) at https://civilrights.osu.edu/.


### Intellectual Diversity

Ohio State is committed to fostering a culture of open inquiry and intellectual diversity within the classroom. This course will cover a range of information and may include discussions or debates about controversial issues, beliefs, or policies. Any such discussions and debates are intended to support understanding of the approved curriculum and relevant course objectives rather than promote any specific point of view. Students will be assessed on principles applicable to the field of study and the content covered in the course. Preparing students for citizenship includes helping them develop critical thinking skills that will allow them to reach their own conclusions regarding complex or controversial matters.

### Religious accommodations
Ohio State has had a longstanding practice of making reasonable academic accommodations for students' religious beliefs and practices in accordance with applicable law. In 2023, Ohio State updated its practice to align with new state legislation. Under this new provision, students must be in early communication with their instructors regarding any known accommodation requests for religious beliefs and practices, providing notice of specific dates for which they request alternative accommodations within 14 days after the first instructional day of the course. Instructors in turn shall not question the sincerity of a student's religious or spiritual belief system in reviewing such requests and shall keep requests for accommodations confidential.

With sufficient notice, instructors will provide students with reasonable alternative accommodations with regard to examinations and other academic requirements with respect to students' sincerely held religious beliefs and practices by allowing up to three absences each semester for the student to attend or participate in religious activities. Examples of religious accommodations can include, but are not limited to, rescheduling an exam, altering the time of a student's presentation, allowing make-up assignments to substitute for missed class work, or flexibility in due dates or research responsibilities. If concerns arise about a requested accommodation, instructors are to consult their tenure initiating unit head for assistance.

A student's request for time off shall be provided if the student's sincerely held religious belief or practice severely affects the student's ability to take an exam or meet an academic requirement and the student has notified their instructor, in writing during the first 14 days after the course begins, of the date of each absence. Although students are required to provide notice within the first 14 days after a course begins, instructors are strongly encouraged to work with the student to provide a reasonable accommodation if a request is made outside the notice period. A student may not be penalized for an absence approved under this policy.

If students have questions or disputes related to academic accommodations, they should contact their course instructor, and then their department or college office. For questions or to report discrimination or harassment based on religion, individuals should contact the Civil Rights Compliance Office.

### Weather or other short-term closing 
Although Ohio State strives to remain open to ensure continuity of services to students and the public, extreme conditions can warrant the usage of the university's Weather or Other Short-Term Closing Policy. Please visit this webpage to learn more about preparing for potential closings and planning ahead for winter weather.


### Counseling and Consultation Services/Mental Health

As a student you may experience a range of issues that can cause barriers to learning, such as strained relationships, increased anxiety, alcohol/drug problems, feeling down, difficulty concentrating and/or lack of motivation. These mental health concerns or stressful events may lead to diminished academic performance or reduce a student's ability to participate in daily activities. The Ohio State University offers services to assist you with addressing these and other concerns you may be experiencing. If you or someone you know are suffering from any of the aforementioned conditions, you can learn more about the broad range of confidential mental health services available on campus via the Office of Student Life Counseling and Consultation Services (CCS) by visiting ccs.osu.edu or calling (614) 292- 5766. CCS is located on the 4th Floor of the Younkin Success Center and 10th Floor of Lincoln Tower. You can reach an on-call counselor when CCS is closed at (614) 292-5766 and 24 hour emergency help is also available through the 24/7 National Suicide Prevention Hotline at 1-(800)-273-TALK or at suicidepreventionlifeline.org

David Wirt, wirt.9@osu.edu, is the CFAES embedded mental health counselor in Columbus. He is available for new consultations and to establish routine care. To schedule with David, please call 614-292-5766. Students should mention their affiliation with CFAES when setting up a phone screening.

Dr. Schaad, schaad.15@osu.edu, is the CFAES embedded mental health counselor in Wooster. She is available for new consultations and to establish routine care. To schedule with 

Dr. Schaad, please call 614-292-5766. Students should mention their affiliation with CFAES when setting up a phone screening.

### Land Acknowledgement

We would like to acknowledge the land that The Ohio State University occupies is the ancestral and contemporary lands of the Shawnee, Potawatomi, Delaware, Miami, Peoria, Seneca, Wyandotte, Ojibwe and Cherokee peoples. The university resides on land ceded in the 1795 Treaty of Greeneville and the forced removal of tribes through the Indian Removal Act of 1830. We honor the resiliency of these tribal nations and recognize the historical contexts that have and continue to affect the Indigenous peoples of this land.

### Accessibility accomodations

The university strives to maintain a healthy and accessible environment to support student learning in and out of the classroom. If you anticipate or experience academic barriers based on your disability (including mental health, chronic, or temporary medical conditions), please let me know immediately so that we can privately discuss options. To establish reasonable accommodations, I may request that you register with Student Life Disability Services. After registration, make arrangements with me as soon as possible to discuss your accommodations so that they may be implemented in a timely fashion.

If you are ill and need to miss class, including if you are staying home and away from others while experiencing symptoms of a viral infection or fever, please let me know immediately. In cases where illness interacts with an underlying medical condition, please consult with Student Life Disability Services to request reasonable accommodations. You can connect with them at slds@osu.edu; 614-292-3307; or slds.osu.edu.

