---
title: "Module 2 Assignment"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is your assignment for Module 2, focused on the material you learned in the lectures and recitation activities on R Markdown, wrangling, ggplot101, and ggplot102. 

Submission info: you will submit this assignment by uploading a knitted `.html` to Carmen. 

- Make sure you include the Code Download button
- Show your code within your knitted `.html`
- Customize the YAML and the document so you like how it looks

Remember there are often many ways to reach the same end product.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(readxl)
library(knitr)
library(tidyverse)
due_dates <- read_excel("../../../files/schedule_modules.xlsx",
                       sheet = "module_due_dates_2026")
```

```{r, echo = FALSE}
module_2_due_date <- due_dates |> 
  filter(Assignment == "Module 2: Coding Fundamentals") |> 
  pull(due_date_paste)
```

::: callout-important
This assignment will be due on **`r module_2_due_date` at 11:59pm**.
:::


