---
title: "Annotating Statistics onto Plots Recitation"
author: "Jessica Cooperstone"
subtitle: "Week 9"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Today you will be practicing what we learned in today's class on adding statistics to your plots.

### Load data

We will be using the NHANES data again from the package [`NHANES`](https://cran.r-project.org/web/packages/NHANES/NHANES.pdf). 

Belwo are the packages that I used to make my plot.

```{r load libraries, message = FALSE, warning = FALSE}
library(tidyverse)
library(NHANES)
library(rstatix)
library(ggpubr)
library(glue)
library(rcompanion)
```

## Is total cholesterol (`TotChol`) different by age (`AgeDecade`)?

::: {.callout-tip collapse="true"}
#### Need a hint? (Click to expand)

Hint - you want to test your assumptions to see what tests to do. You might need to use different posthoc comparison methods than we did in class.

:::

::: {.callout-tip collapse="true"}
#### Need another hint? (Click to expand)

Another hint - the function [`rcompanion::cldList()`](https://www.rdocumentation.org/packages/rcompanion/versions/2.4.30/topics/cldList) will convert the resulting comparison table from a posthoc Dunn test to create a column with the letters indicating which groups are significantly different from each other.

:::